### 2021-05-18 chatbot_classification/03.emnlp_2021/experiments.tex
modified:   experiments.tex   modified:   methodology.tex
  new file:   paper.ver03.pdf

All pretrained model baselines are fine-tuned on our datasets.
Especially, the similarity model baseline and all similarity layers in SFCs use extra Quora dataset \cite{iyer2017first} to enhance system performance. 
This is also one of merits in SFC, as it support adopting out-of-domain data in comparison to a classification based model.

>>>>

All pretrained model baselines are fine-tuned on our datasets.
Especially, the similarity model baseline and all similarity layers in SFCs use extra Quora dataset \cite{iyer2017first} \footnote{which contains 404,290 potential duplicate question pairs} to enhance system performance. 
This is also one of merits in SFC, as it support adopting out-of-domain data in comparison to a classification based model.

### 2021-05-18 chatbot_classification/03.emnlp_2021/methodology.tex
modified:   experiments.tex   modified:   methodology.tex
modified:   experiments.tex   modified:   methodology.tex
  new file:   paper.ver03.pdf

\section{Methodology}
In  this section, we describe our proposed SFC method, which includes the basic 2-stage SFC and its more compact version joint-SFC.

@@

It consists of the following  two  independent stages  in  charge of different  jobs.

\subsubsection*{
  Stage 1: classification model for providing top-K candidate class labels
} 
We  can  use any auxiliary text classification model or tool,
such as Term Frequency-Inverse Document Frequency (TF-IDF)retrieval, to
provide  top-K  most  related  class  labels. These selected class labels will
later  be  used to sample sentence pairs of both positive and negative
samples  for  stage  2.

In  this  work,  we  choose pretrained models to fine tune on our data.
Due to the excellent performance of RoBERTa in this task, compared to other pretrained
models, we use RoBERTa as the encoder of classification model in our setting.

Given   a   data   point   $x_{i}$  with  class  label  $y_{i}$  from  dataset
$\mathcal{D}$, we take the final hidden state $h_{i}$ of the first token [CLS]
encoded  by  RoBERTa  as the representation for the whole sequence of $x_{i}$.
Then  a linear layer is followed to output probabilistic distribution of class
labels,  $softmax(W^Ch_{i}+b^C)=softmax({\Phi}^C_{i})$,  where $W^C$ and $b^C$
are  trainable  parameters.  Then, the loss in stage 1 classification model is
shown in Equ. \ref{eq:classification_loss},

\vspace{-2em}
\begin{equation}
  \begin{aligned}
    \mathcal{L}^{C}&=\frac{1}{N}\sum_{i=1}^{N}-log(\frac{exp(\varPhi_{i,y_{i}}^C)}{\sum_{j=1}^{C}exp(\varPhi_{i,j}^C)}) \\
    \label{eq:classification_loss}
  \end{aligned}
\end{equation}
\vspace{-2.5em}

\subsubsection*{
  Stage 2: sentence-pair similarity model based classification with multi-task
  learning
}
We  continue choosing RoBERTa as the main module in this stage that identifies
the class label with strongest semantic similarity to the user input sentence.

Suppose  we have a training dataset $\mathcal{D}=\{x_{i},y_{i}\}_{i=1}^{N}$ of
$N$  data  points,  in  which $x_{i}$ is the user input query and $y_{i}$ is a
single  class label from a class label set of $C$ classes in total. We 
generate            a sentence            pair            dataset
$\mathcal{D'}=\{[(x,x')_{j},l_{j}]\}_{j=1}^{M}$ from $\mathcal{D}$,   where   $(x,   x')_{j}$  is
a pair of two history queries \footnote{For simplicity, we use query
to denote a user's question or a response.}, and $l_{j}\in\{0, 1\}$ denotes the
similarity label that only the ones from the same class label are considered
similar. The size of $\mathcal{D'}$  would be up to $O(N^2)$ if no any
sampling strategy is used.  This is quite time-consuming  for the  training
in stage 2 because $N$ can be several hundreds or thousands even under 
few shot setting in task-specific chatbot scenario.

Therefore,  we  adopt  the idea of adversarial negative sampling strategy from
the  work  \cite{bamler2020extreme}.  The  key  idea is to train with negative
samples  that  are hard to be distinguished from positive samples. Now suppose
we  have  a  user  input query $x_{i}$ and top-K most related class labels set
$\mathcal{C'}=\{c'_{1},  \dots,  c'_{K}\}$ provided by auxiliary model or tool
in  stage  1, where $K$ is a hyperparameter to control candidate class number,
we  apply  the  adversarial  negative sampling strategy into our sentence pair
sampling process by the following two steps:

The  first step is \emph{positive sentence pair sampling}. During the training
process,  we  first make sure the ground truth class label $y_{i}$ for $x_{i}$
is  within  the  candidate  class  label  set  $\mathcal{C'}$. If $y_{i}\notin
\mathcal{C'}$,  we  manually  add  $y_{i}$  into  $\mathcal{C'}$  by replacing
$c'_{K}$  with  $y_{i}$, since $c'_{K}$ is the least promising candidate label
according  to  the  auxiliary  model  in stage 1. Afterwards, we will randomly
sample  $P$ sentences $\{x'_{1}, \dots, x'_{P}\}$ with the same class label as
$y_{i}$  from  dataset  $\mathcal{D}$  to  form the set of sentence pairs with
positive  label  $\mathcal{P'}_{i}=\{[(x_{i},  x'_{1}),  1],  \dots,  [(x_{i},
x'_{P}),  1]\}$,  where $P$ is also a hyperparameter set to control the number
of sentences we should sample from each class.

\begin{table*}
  \begin{centering}
    \scalebox{0.75}{
      \begin{tabular}{|c|c|c|c|c|c|}
        \hline 
        \textbf{Dataset} & \textbf{Domain} & \textbf{\#Class} &
        \textbf{\#Training Samples} & \textbf{\#Samples/Class} & \textbf{Settings}\tabularnewline
        \hline 
        ITG & FAQ chatbot & 228 & 3938 * 0.7 & 12 & 3-fold\tabularnewline
        \hline 
        Amazon-670k & Product review & 250 & 2658 * 0.7 & 8 & 3-fold\tabularnewline
        \hline 
        HWU64 & Intent detection & 64 & [320, 640, 960, 1280, 1920, 3200] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
        CLINC150 & Intent detection & 150 & [750, 1500, 2250, 3000, 4500, 7500] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
        BANKING77 & Intent detection & 77 & [385, 770, 1155, 1540, 2310, 3850] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
        FRAMES & Intent detection with NER & 21 & [208, 408, 984] & [10, 20, 50] & data sampling\tabularnewline
        \hline
        ATIS & Intent detection with NER & 21 & [168, 303, 544] & [10, 20, 50] & data sampling\tabularnewline
        \hline
      \end{tabular}
    }
    \par
  \end{centering}
  \caption{Statistics for all datasets and few shot settings.}

  \label{tbe:dataset statistic}
\end{table*}

The  second  step  is  \emph{negative sentence pair sampling}: As for negative
sentence  pairs,  we will also randomly sample $P$ sentences for each class in
the negative candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$. The
class  labels  in  $\mathcal{C'}\backslash  \{y_{i}\}$  are  the  set  of most
confusing  class  labels comparing to the ground truth ${y_{i}}$, so we assume
that  the  sentence  pairs  with  negative  label  grouped by user input query
$x_{i}$  and sentences with class labels in $\mathcal{C'}\backslash \{y_{i}\}$
are  strong  adversarial  negative  samples for sentence-pair similarity model
that  can  help  enhance  the training speed and performance. According to the
same  method  as  positive  sentence  pair  sampling, we can obtain the set of
sentence  pairs  with  negative label as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{1}),
0], ..., [(x_{i}, x'_{P\cdot K}), 0]\}$

In  this way, we generate the sentence pair dataset $\mathcal{D'}$ for stage 2
based  on  the  top-K  class  label  set $\mathcal{C'}$ provided by stage 1 as
$\mathcal{D'}=\{\mathcal{P'}_{i}\cup  \mathcal{N'}_{i}\}_{i=1}^{N}$,  in which
we have $M=N\cdot K\cdot P$ data points of sentence pairs in total.

Before starting to fine-tune our sentence-pair model on task-specific dataset,
we  first  fine-tune  RoBERTa on Quora dataset  \cite{iyer2017first}, which
contains  404,290  potential  duplicate question pairs, for transfer learning.
This  is  also  the  merit of similarity based model, as labeled task-specific
classification  data  is  hard  to  acquire,  yet  general sentence pairs with
similar semantics can be much easier to acquire.

\subsubsection*{Multi-task based training}
We   continue  to  do  multi-task  training  on  sentence-pair  dataset
$\mathcal{D'}$ sampled from the top-K candidate class labels provided by stage
1. In stage 2, we have two tasks tuning our system.

The first task is \emph{regular sentence-pair similarity task}.
For  sentence  pair semantic similarity task, given a data point $(x, x')_{i}$
with       similarity       label       $l_{i}$       from       data      set
$\mathcal{D'}=\{[(x,x')_{i},l_{i}]\}_{i=1}^{M}$,   Roberta   takes  the  final
hidden  state  $h_{i}$  of the first token [CLS] as the representation for the
sequence of packed sentence pair $(x, x')_{i}$. Let's suppose we have a linear
layer  as  ${\Phi}^S_{i}=W^Sh_{i}+b^S$,  where  $W^S$  and $b^S$ are trainable
parameters,  and  the probability score for $(x, x')_{i}$ can be calculated as
$p^S_{i}=softmax({\Phi}^S_{i})$.

Due  to  the  fact that sentence pairs within the same class is much fewer than
that from  different  classes,  the  dataset  $\mathcal{D'}$  is
quite imbalanced.  Therefore, we will accommodate a weight variable $w^S = [K-1, 1]$
to  the  loss  to eliminate the bias brought by data imbalance, shown in Equ.
\ref{eq:similarity_loss}.

@@

The  second  task  is  \emph{classification  on  top-K classes}. As minimizing
similarity  loss  is  not  our  final goal in the intent classification, we
bring  back  classification  loss again. In this task, the network is based on
sentence  pair  similarity  modules.  The  only  difference  is  that we add a
task-specific  average  pooling  layer,  shown  in Fig. \ref{fig:framework} to
accomplish the classification task based on sentence-pair model.

We  already  know  that the size of ${\Phi}^S$ in task 1 is $M\times 2=(N\cdot
K\cdot  P)\times  2$, where $N$ is the total number of data points in original
training  data $\mathcal{D}$, $K$ is a hyperparameter that controls the number
of   candidate   class  labels  provided  by  stage  1,  and  $P$  is  also  a
hyperparameter that controls the number of sentences we should randomly sample
from each candidate class labels. Now, for the average pooling layer, we first
reshape  ${\Phi}^S$  into  the  size  of $N\times K\times P\times 2$, and then
split  it  into  ${\Phi}^{K,pos}$  and ${\Phi}^{K,neg}$, and both of them will
have  the  size  of  $N\times  K\times  P$.  In this way, ${\Phi}^{K,pos}$ can
represent  the  level  of  similarity  for each sentence pair, and in the mean
time,  ${\Phi}^{K,neg}$  can  represent  the  level  of dissimilarity for each
sentence  pair. Then, we can do average pooling for each candidate class among
the  top-K  classes  as  shown  in  Equ.  \ref{eq:average  pooling}.

@@

\subsection{Joint-SFC}
In  2-stage  SFC,  Stage  1  and  stage 2 are separate that there is no deep
interaction  with each other during training. In this case, the performance of
stage  1  may  limit  the potential of stage 2; meanwhile, stage 2 also cannot
give  training feedback back to stage 1 for fine-tuning. Therefore, to further
improve  the  overall performance of SFC, we proposed a joint model structure,
shown in Fig. \ref{fig:framework}.

>>>>

\section{Methodology}
In this section, we describe our proposed SFC method, which includes a basic 2-stage SFC implementation and a more compact version called joint SFC.

@@

We first introduce this most intuitive implementation.
The outputs of the first stage are fed into the second stage, yet its parameters can not be tuned based on feedback of the second stage.

\subsubsection*{Stage 1: top-K candidate class labels generation} 
Any model that generates top-K most related class labels can be adopted in stage 1, such as classification models, term frequency-inverse document Frequency (TF-IDF) based retrieval models.
Due to the excellent performance of RoBERTa in practice, we use RoBERTa to construct a classification model in this stage.

Suppose we have a training dataset $\mathcal{D}=\{x_{i},y_{i}\}_{i=1}^{N}$, in which $x_{i}$ is the user input query and $y_{i}$ is a single class label in $\{1, \ldots, J\}$. 
Given $x_i$, we take the final hidden state $h_i$ of the first token [CLS] encoded by RoBERTa as the representation for the whole sequence of $x_i$.
A linear layer is followed to output unnormalized probabilistic distribution of class labels, ${\Phi}_i^C=W^C h_i + b^C$, where $W^C$ and $b^C$ are class label related trainable parameters. 
Then, the loss in stage 1 classification model is shown in Equ. \ref{eq:classification_loss},

\vspace{-2em}
\begin{equation}
  \begin{aligned}
    \mathcal{L}^{C}&=\frac{1}{N} \sum_{i=1}^{N}-log(\frac{exp(\varPhi_{i,y_{i}}^C)}{\sum_{j=1}^{J}exp(\varPhi_{i,j}^C)}) \\
    \label{eq:classification_loss}
  \end{aligned}
\end{equation}
\vspace{-2.5em}

\subsubsection*{Stage 2: sentence-pair similarity model based multi-task learning}
We continue choosing RoBERTa to construct main modules in this stage.

We generate a sentence pair dataset $\mathcal{D'}=\{[(x,x')_{j},l_{j}]\}_{j=1}^{M}$ from $\mathcal{D}$, where $(x, x')_{j}$ is a pair of two history queries \footnote{For simplicity, we use query to denote a user's question or a response.}, and $l_{j}\in\{0, 1\}$ denotes the similarity that only the pair from the same class label is considered similar. 
The size of $\mathcal{D'}$ would be up to $N^2$ without a sampling strategy. 
This is prohibitively large and time-consuming for model training in stage 2, because $N$ can be several hundreds or thousands even under few shot setting in our scenario.

Therefore, we adopt the idea of adversarial negative sampling strategy from the work \cite{bamler2020extreme}. 
The key idea is to train with negative samples that are hard to be distinguished from positive samples. 
Now suppose we have a user input query $x_{i}$ and top-$K$ most related class label set $\mathcal{C'}=\{c'_{1}, \dots, c'_{K}\}$ from stage 1, where $K$ is a hyperparameter, we apply the adversarial negative sampling strategy into our sentence pair sampling in two steps:

The first step is \emph{positive sentence pair sampling}. 
In training, we make sure the ground truth class label $y_{i}$ of $x_{i}$ exists in the candidate class label set $\mathcal{C'}$. 
Otherwise, we have it substitute for the least relevant class label $c'_{K}$. 
Then, we randomly sample $P$ sentences $\{x'_{1}, \dots, x'_{P}\}$ with the same class label as $y_{i}$ from dataset $\mathcal{D}$ to form the set of sentence pairs with positive label $\mathcal{P'}_{i}=\{[(x_{i}, x'_{1}), 1], \dots, [(x_{i}, x'_{P}), 1]\}$, where $P$ is also a hyperparameter.

The second step is \emph{negative sentence pair sampling}.
We also randomly sample $P$ sentences for each class in the negative candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$, which is the more confusing to the expected class label compared to those out of top-$K$ candidate set.
The resulting sentence pairs are strong adversarial negative samples for sentence-pair similarity model, and they significantly enhance the training speed and performance. 
We denote them as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{1}), 0], ..., [(x_{i}, x'_{P\cdot K}), 0]\}$

In this way, we generate the sentence pair dataset $\mathcal{D'}$ for stage 2, based on the top-K class label set $\mathcal{C'}$ from by stage 1, as $\mathcal{D'}=\{\mathcal{P'}_{i}\cup \mathcal{N'}_{i}\}_{i=1}^{N}$, and its size is $M=N\times K\times P$.

As analyzed before, one advantage of similarity based model is its potential to use out-of-domain data. We take this advantage and use extra Quora dataset \cite{iyer2017first} to enhance performance of both baseline and our SFCs.

\subsubsection*{Multi-task based training}
We have our stage 2 model train on two tasks in the framework of multi-task training.

The first task is \emph{regular sentence-pair similarity task}.
For sentence pair semantic similarity task, given a data point $(x, x')_{i}$ with similarity label $l_{i}$ from data set $\mathcal{D'}$, Roberta takes the concatenation of $x, x'$ and outputs hidden state $h_{i}$ of the first token [CLS] as its representation.
A linear layer is followed as ${\Phi}^S_{i}=W^Sh_{i}+b^S$ to generate unnormalized probabilities, where $W^S$ and $b^S$ are trainable parameters.

We note that the dataset $\mathcal{D'}$ is unbalanced, as negative samples are $K - 1$ times as many as positive samples.
Therefore, we add a weight $w^S \in \{K-1, 1\} $ to the loss of positive and negative samples respectively to eliminate the bias, shown in Equ. \ref{eq:similarity_loss}.
This weighting strategy is quite effective in practice.

@@

The second task is \emph{classification on top-K classes}. 
As minimizing similarity loss is not our final goal of the chatbot application, we bring back classification loss again. 
In this task, the network structure is based on sentence pair similarity modules. 
We then add a task-specific average pooling layer, shown in Fig. \ref{fig:framework} to accomplish the classification task based on sentence-pair model.

The size of ${\Phi}^S$ in task 1, namely unnormalized probabilistic distribution, is $2 \times M$, as $M$ sentence pairs corresponds two class labels.
Further, it equals $(N\cdot K\cdot P)\times 2$, where $N$ is the total number of data points in original training data $\mathcal{D}$, $K$ is a hyperparameter that controls the number
of candidate class labels provided by stage 1, and $P$ is also a hyperparameter that controls the number of sentences we should randomly sample from each candidate class labels. 
Now, for the average pooling layer, we first reshape ${\Phi}^S$ into the size of $N\times K\times P\times 2$, and then split it into ${\Phi}^{K,pos}$ and ${\Phi}^{K,neg}$, and both of them will have the size of $N\times K\times P$. 
In this way, ${\Phi}^{K,pos}$ represents the level of similarity for each sentence pair, and ${\Phi}^{K,neg}$ represents the level of dissimilarity for each sentence pair. 
Then, we can do average pooling for each candidate class among the top-K classes as shown in Equ. \ref{eq:average pooling}.

@@

\subsection{Joint SFC}
In 2-stage SFC, Stage 1 and stage 2 are separate that there is no deep
interaction with each other during training. In this case, the performance of
stage 1 may limit the potential of stage 2; meanwhile, stage 2 also cannot
give training feedback back to stage 1 for fine-tuning. Therefore, to further
improve the overall performance of SFC, we proposed a joint model structure,
shown in Fig. \ref{fig:framework}.
